---
title: "Final Exam Question 3"
author: "Karmanya Pathak"
date: "4/25/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


(1) ANOVA:

Capture an ANOVA model on the iris dataset. Set the dependent variable to 'Species'.
Capture a summary. 
Provide an interpretation of the results in your own words.  Support your response with results captured from running ANOVA.


- id: ID
- Species: 3 Species of flowers
- Sepal.Length: Sepal length in cms
- Sepal.width: Sepal width in cms
- Petal.Length: Petal length in cms
- Petal.Width: Petal width in cms

```{r iris}

iris_data <- read.csv("D:/Harrisburg/Lectures/Spring 2020/ANLY 500 - Principles of Analytics I/Final Exam/Question 3/iris_exams.csv")

#Using a copy of the dataset in case we need to access the original data

iris_q3 <- iris_data

summary(iris_q3)


```



```{r library}

#Loading all the libraries

library(MOTE)
library(reshape)
library(pwr)
library(ggplot2)
library(corrplot)
library(moments)
library(biotools)
library(ez)
library(biotools)
library(knitr)
library(kableExtra)


#Use the template for the plots

cleanup = theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(), 
                panel.background = element_blank(), 
                axis.line.x = element_line(color = "black"),
                axis.line.y = element_line(color = "black"),
                legend.key = element_rect(fill = "white"),
                text = element_text(size = 15))
  



```


## Data Screening:

## Accuracy, Missing Data:

The data is clean and no missing data (NAs).

Since there are no missing values, there is no need to omit any observations.

```{r Data Screening}

#Accuracy, missing data

summary(iris_q3)



```


## Outliers:

The cut off score for the mahalanobis measure is 18.46683

There were 2 outliers in the dataset


```{r Outliers}

# Since we can calculatate means of variables using mahalanobis(in general as well) on only numeric variables, 
# we would be ignoring factor variables id and Species

mahal = mahalanobis(iris_q3[,-c(1,2)], colMeans(iris_q3[,-c(1,2)]), cov(iris_q3[,-c(1,2)]))

#Calculating the cutoff score

cutoff = qchisq(0.999, ncol(iris_q3[,-c(1,2)]))


#Degrees of freedom

#ncol(iris[,-c(1,2)])

#Printing the cutoff score

cutoff

#Checking to see number of outliers in iris

summary(mahal > cutoff)

#Creating a new dataframe without the 2 outliers

iris_no_outliers_q3 <- subset(iris_q3, mahal < cutoff)

#Checking the new dataframe without the outliers

str(iris_no_outliers_q3)

summary(iris_no_outliers_q3)


```


## Assumptions:

## Additivity:

The model meets the assumption for additivity, and we have multicollinearity.

```{r }
iris_numeric <- iris_no_outliers_q3[,c(3:6)]


corrplot(cor(iris_numeric))

cor(iris_numeric)


```

## Linearity: 

I think the model meets the assumption for linearity.

We are deviating from linearity, but there are not enough points to dismiss or violate linearity.

```{r linearity}

#We are removing one factor variable id to run a linear model

iris_no_id_q3 <- iris_q3[,-1]


random <- rchisq(nrow(iris_no_id_q3),4)
fake <- lm(random~., data = iris_no_id_q3)
#summary(fake)
standardized <- rstudent(fake)
fitted <- scale(fake$fitted.values)

#Generating the linearity graph

{qqnorm(standardized)
abline(0,1)}




```


## Normality: 


Although we have a slight skew in the data, it is not enough to violate normality.


```{r normality}


hist(standardized, breaks=15)


```


## Homogeneity/Homoscedasticity: 

# Homogeneity:

Using Box's M test, we find that p < 0.001, stating that we haven't met the assumption for Homogeneity.


#Homoscedasticity:

Apart from a few observations, the scatter is fairly equal from the lowest point to the largest point, letting us accept the assumption for homoscedasticity.




```{r homog-s}


#Plotting the values:

plot(fitted, standardized) 
abline(0,0)
abline(v = 0)

#Using Box's test to check for Homogeneity

boxM(iris_no_id_q3[,-1], iris_no_id_q3[,1])




```


## Analysis of Variance

Capture an ANOVA model on the iris dataset. Set the dependent variable to 'Species'.
Capture a summary. 
Provide an interpretation of the results in your own words.  Support your response with results captured from running ANOVA.

The omnibus ANOVA test was significant, since p value is low (<0.05)




```{r AOV}

##you must have a participant number for ezANOVA
iris_no_outliers_q3$partno = 1:nrow(iris_no_outliers_q3)

options(scipen = 999)
ezANOVA(data = iris_no_outliers_q3,
        dv = Sepal.Length,
        between = Species,
        wid = partno,
        type = 3, 
        detailed = T)

# Since Homogeneity was bad

oneway.test(Sepal.Length ~ Species, data = iris_no_outliers_q3)


aovModel1 <- ezANOVA(data = iris_no_outliers_q3,
        dv = Sepal.Length,
        between = Species,
        wid = partno,
        type = 3, 
        detailed = T)


# Coeeficient covariances 

aovModel1


```


## Effect Sizes:

The effect size is large

```{r effect sizes}

eta <- aovModel1$ANOVA$ges[2]

eta

## formula = (SSM - dfM*MSR)/ (SStotal + MSR)

MSR <- aovModel1$ANOVA$SSd[2] / aovModel1$ANOVA$DFd[2]

w2 <- (aovModel1$ANOVA$SSn[2] - aovModel1$ANOVA$DFn[2]*MSR) / (aovModel1$ANOVA$SSn[2] + aovModel1$ANOVA$SSd[2] + MSR)
w2

```

## Post Hoc t-test


```{r posthoc}

pairwise.t.test(iris_no_outliers_q3$Sepal.Length,
                iris_no_outliers_q3$Species,
                p.adjust.method = "none", 
                paired = F, 
                var.equal = T)

pairwise.t.test(iris_no_outliers_q3$Sepal.Length,
                iris_no_outliers_q3$Species,
                p.adjust.method = "bonferroni", 
                paired = F, 
                var.equal = T)

#Unequal variances

post1 <- pairwise.t.test(iris_no_outliers_q3$Sepal.Length,
                iris_no_outliers_q3$Species,
                p.adjust.method = "none", 
                paired = F, 
                var.equal = F)

post2 <- pairwise.t.test(iris_no_outliers_q3$Sepal.Length,
                iris_no_outliers_q3$Species,
                p.adjust.method = "bonferroni", 
                paired = F, 
                var.equal = F)

```


#Effect Sizes:

Using d from the ANOVA test to calculate effect sizes


```{r effectsize}

M <- with(iris_no_outliers_q3, tapply(Sepal.Length, Species, mean))
stdev <- with(iris_no_outliers_q3, tapply(Sepal.Length, Species, sd))
N <- with(iris_no_outliers_q3, tapply(Sepal.Length, Species, length))

## Setosa vs Versicolor is 1 and 2
effect1 = d.ind.t(m1 = M[1], m2 = M[2],        
sd1 = stdev[1], sd2 = stdev[2],        
n1 = N[1], n2 = N[2], a = .05)
effect1$d



```

# Creating a tabular view by Type of Post Hoc:

```{r table, results='asis'}


tableprint = matrix(NA, nrow = 3, ncol = 2)

##row 1
tableprint[1, ] = c("No correction p", post1$p.value[1])

##row 2
tableprint[2, ] = c("Bonferroni p", post2$p.value[1])

##row 3
tableprint[3, ] = c("d value", effect1$d)

#don't change this
kable(tableprint, 
      digits = 3,
      col.names = c("Type of Post Hoc", 
                    "Setosa vs Versicolor"))
```


#Running a trend analysis:

This is a significant trend since the p value is low, we have a high adjusted R-squared value and the significant code values are 0 and 0.001 implying high significance.


```{r trend}
#Trend Analysis

k = 3
iris_no_outliers_q3$part <- iris_no_outliers_q3$Species
contrasts(iris_no_outliers_q3$part) = contr.poly(k)
output2 = aov(Sepal.Length ~ part, data = iris_no_outliers_q3)
summary.lm(output2)


```

#Bar Chart

The Sepal length of Virginica species is comparatively higher than the other two species.

```{r graph}

bargraph = ggplot(iris_no_outliers_q3, aes(Species, Sepal.Length))
bargraph +
  cleanup +
  stat_summary(fun = mean, 
               geom = "bar", 
               fill = "White", 
               color = "Black") +
  stat_summary(fun.data = mean_cl_normal, 
               geom = "errorbar", 
               width = .2, 
               position = "dodge") +
  xlab("Species") +
  ylab("Sepal Length") +
  coord_cartesian(ylim = c(0,7)) +
  scale_x_discrete(labels = c("Setosa","Versicolor","Virginica"))



```







