---
title: "Introduction to Data Analytics 1"
author: "Karmanya Pathak"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Variables, Hypothesis, Designs

*Title:* Offshore outsourcing: Its advantages, disadvantages, and effect on the American economy

*Abstract*: The United States has trained some of the world's best computer programmers and technology experts. Despite all of this training, many businesses do not have a full understanding of information technology. As the importance of technology in the business world grows, many companies are wasting money on extensive technology projects. When problems arise, they expect that further investment will solve these issues. To prevent such problems, many companies have begun to outsource these functions in an effort to reduce costs and improve performance. The majority of these outsourced information technology and call center jobs are going to low-wage countries, such as India and China where English-speaking college graduates are being hired at substantially lower wages. The purpose of this study is to evaluate the positive and negative aspects of offshore outsourcing with a focus on the outsourcing markets in India and China, arguably the two most popular destinations for outsourcers. The cost savings associated with offshore outsourcing will be evaluated in relation to the security risks and other weakness of offshore outsourcing. In addition, an analysis of the number of jobs sent overseas versus the number of jobs created in the United States will be used to assess the effects that outsourcing is having on the American economy and job market. Finally, the value of jobs lost from the American economy will be compared to the value of jobs created. The goal of these analyses is to create a clear picture of this increasingly popular business strategy.

Answer the following questions about the abstract above:

1)	What is a potential hypothesis of the researchers?

Offshore outsourcing technological functions to the two most popular outsourcing market destinations, India and China, is an increasingly popular business strategy.

2)	What is one of the independent variables?

where - (destination/location) is one of the independent variables
  
    a.	What type of variable is the independent variable?

It is a variable that is the proposed cause and that which can be used as a predictor variable. It is a variable that stands alone and does not change with the other variables we are trying to measure, i.e., it has no effect on itself when any other variable changes. Also, referred from the lecture notes, it is a manipulated variable in experiments and has valu(s) in hypothesis.

3)	What is one of the dependent variables?

cost and cost2 are examples of dependent variables.
  
    a.	What type of variable is the dependent variable?

It is a variable that is the proposed effect and that which can be used as a outcome variable. It is a variable that depends on the independent variable, i.e., a change in any one or all the independent variables can have an effect on dependent variables. Referrencing the notes, it is a measures and not a manipulated variable in experiments and has value in hypothesis.


4)	What might cause some measurement error in this experiment?

The discrepancy in the cost of jobs and contracts with offshore markets due to volatility in global financial markets after signing the contract/agreement may cause some measurement error in the total cost during the length of the contract/job.


5)	What type of research design is the experiment?

The research design of this experiment is based on correlational research,

    a.	Why?

Cost and number of jobs is related to the location of the project.

6)	How might you measure the reliability of your dependent variable?

We can measure the reliability of our dependent variable by undergoing test-retest reliability measurement.
Both intra-rater reliability and inter-rater reliability forms can be used to measure the reliability of our dependent variable in this experiment.

7)	Is this study ecologically valid?
This study is ecologically valid in the sense that it can be valid as a real-world problem.

8)	Can this study claim cause and effect?
    Yes, this study can claim cause and effect.

    a.	Why/why not?
The location of the job affects the number of jobs and the cost of each job as well.

9)	What type of data collection did the researchers use (please note that #5 is a different question)?
The data collection that the researchers used appears to be between-groups/independent and it is evident that different entities are used in this experimental condition.


# Part 2: Use the assessment scores dataset (03_lab.csv) to answer these questions.

The provided dataset includes the following information created to match the abstract:

- Jobs: the percent of outsourced jobs for a call center.
- Cost: one calculation of the cost savings for the business.
- Cost2: a separate way to calculate cost savings for the business.
- ID: an ID number for each business.
- Where: where the jobs were outsourced to. 

Calculate the following information:

1)	Create a frequency table of the percent of outsourced jobs.

```{r}
df <- read.csv("C:\\Users\\Karmanya\\Desktop\\Harrisburg\\Lectures\\Spring 2020\\ANLY 500 - Principles of Analytics I\\Assignment 2\\03_data.csv")
table(df$jobs)
subset(df,select = 'jobs')
library(dplyr)
res <- df %>% group_by(jobs) %>% summarise(Freq=n())
res
```

2)	Create histograms of the two types of cost savings. You will want to add the breaks argument to the hist() function. This argument adds more bars to the histogram, which makes it easier to answer the following questions:

`hist(dataset$column, breaks = 15)` 

15 is a great number to pick, but it can be any number. For this assignment, try 15 to see a medium number of bars. 

```{r}
hist(df$cost, breaks = 15)
hist(df$cost2, breaks = 15)
```

3)	Examine these histograms to answer the following questions:

    a.	Which cost savings appears the most normal?
    cost2 appears to be the most normal
    b.	Which cost savings data is multimodal?
    cost is multimodal
    c.	Which cost savings data looks the most skewed (and in which direction positive or negative)?  
    cost looks most skewed in the positive direction.
    d.	Which cost savings data looks the most kurtotic?
    cost2 looks the most kurtotic
4)	Calculate the z-scores for each cost savings, so they are all on the same scale.

```{r}
df$sc_cost <- scale(df$cost)

df$sc_cost2 <- scale(df$cost2)

df$sc_cost
df$sc_cost2
```

6)	How many of the cost saving scores were more extreme than 95% of the data (i.e., number of z-scores at a *p* < .05)?

```{r}
df$lg_cost <- pnorm(-abs(df$sc_cost))
df$lg_cost2 <- pnorm(-abs(df$sc_cost2))

nrow(subset(df$sc_cost,df$lg_cost<0.05))
nrow(subset(df$sc_cost2,df$lg_cost2<0.05))


df[which.max(df$cost),]
df[which.max(df$cost2),]

df[which.min(df$cost),]
df[which.min(df$cost2),]

df[which.max(df$sc_cost),]
df[which.max(df$sc_cost2),]

df[which.min(df$sc_cost),]
df[which.min(df$sc_cost2),]
?apply()
```

    a.	Cost Savings 1: 21
    
    c.	Cost Savings 2: 19
    
7)	Which business had:

    a.	the highest cost savings? 
    
    S100 and S97 have the highest cost savings
    
    b.	the the lowest cost savings? 
    
    S190 and S92 have the lowestcost savings
    
    c.  Use both cost savings columns and find the ID number of the business with the lowest and highest z-score.
    
    S100, S97 have the highest z-scores while S190 andS92 have the lowest z-scores
